A White Noise Approach to Evolutionary Ecology

### 2020-04-28 Current Version/draft.Rmd
reorganize

---
title: "A White Noise Approach to Evolutionary Ecology"
author: "Bob Week, Steve Krone, Scott L. Nuismer, Luke J. Harmon"
date: "3/1/2020"
abstract: "We derive the dynamics of the distribution of a quantitative character and the abundance of a biological population from a stochastic partial differential equation driven by space-time white noise. In the process we develop a useful set of heuristics to operationalize the powerful, but abstract theory of white noise and measure-valued stochastic processes. This approach allows us to compute the full implications of demographic stochasticity on phenotypic distributions and abundances of populations. We demonstrate the utility of our approach by deriving a quantitative genetic model of diffuse coevolution mediated by exploitative competition for a continuum of resources. Other than trait and abundance distributions, this model predicts interaction networks parameterized by rates of interactions, competition coefficients, and selection gradients. We briefly investigate the relationship between selection gradients and competition coefficients. This illustrative investigation suggests absolute values of linear selection gradients are negatively correlated with competition coefficients when the ratio of interspecific trait variation to intraspecific trait variation is approximately less than $1.25$ and positively correlated otherwise. The inverse relationship holds for quadratic selection gradients except with threshold diversity ratio equal to unity instead of $1.25$. Hence, this approach contributes to the development of a synthetic theory of evolutionary ecology by formalizing first principle derivations of stochastic models that underlie rigorous investigations of the relationship between feedbacks of biological processes and the patterns of diversity they produce."
bibliography: ref.bib
output:
  pdf_document: 
    extra_dependencies:
      - float
      - flafter
    fig_caption: yes
    fig_height: 3
    fig_width: 4
    keep_tex: yes
    number_sections: yes
header-includes:
  - \usepackage{xr} # for cross-referencing between documents
  - \usepackage{mathtools}
  - \usepackage{mathrsfs}
  - \usepackage{hyperref}
  - \usepackage[left]{lineno}
  - \usepackage{amsmath}
  - \linenumbers
  - \externaldocument{SM} # externel document to cross-reference
editor_options: 
  chunk_output_type: console
---

>>>>

---
title: A White Noise Approach to Evolutionary Ecology
author:
  - name: Bob Week
    email: bobweek@gmail.com
    affiliation: University of Idaho
    footnote: 1
  - name: Steve Krone
    affiliation: University of Idaho
  - name: Scott L. Nuismer
    affiliation: University of Idaho
  - name: Luke J. Harmon
    affiliation: University of Idaho
address:
  - code: University of Idaho
    address: Department of Biological Sciences, 875 Perimeter Drive MS 3051, Moscow, Idaho, 83843
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"
date: "`r Sys.Date()`"
abstract: |
  We derive the dynamics of the distribution of a quantitative character and the abundance of a biological population from a stochastic partial differential equation driven by space-time white noise. In the process we develop a useful set of heuristics to operationalize the powerful, but abstract theory of white noise and measure-valued stochastic processes. This approach allows us to compute the full implications of demographic stochasticity on phenotypic distributions and abundances of populations. We demonstrate the utility of our approach by deriving a quantitative genetic model of diffuse coevolution mediated by exploitative competition for a continuum of resources. Other than trait and abundance distributions, this model predicts interaction networks parameterized by rates of interactions, competition coefficients, and selection gradients. We briefly investigate the relationship between selection gradients and competition coefficients. This illustrative investigation suggests absolute values of linear selection gradients are negatively correlated with competition coefficients when the ratio of interspecific trait variation to intraspecific trait variation is approximately less than $1.25$ and positively correlated otherwise. The inverse relationship holds for quadratic selection gradients except with threshold diversity ratio equal to unity instead of $1.25$. Hence, this approach contributes to the development of a synthetic theory of evolutionary ecology by formalizing first principle derivations of stochastic models that underlie rigorous investigations of the relationship between feedbacks of biological processes and the patterns of diversity they produce.
bibliography: ref.bib
output:
  pdf_document: 
    extra_dependencies:
      - float
      - flafter
    fig_caption: yes
    fig_height: 3
    fig_width: 4
    keep_tex: yes
    number_sections: yes
header-includes:
  - \usepackage{xr} # for cross-referencing between documents
  - \usepackage{mathtools}
  - \usepackage{mathrsfs}
  - \usepackage{hyperref}
  - \usepackage[left]{lineno}
  - \usepackage{amsmath}
  - \usepackage[authoryear,sectionbib,sort]{natbib}  
  - \usepackage{fullpage}
  - \linenumbers
  - \externaldocument{SM} # externel document to cross-reference
editor_options: 
  chunk_output_type: console
---

### 2020-04-28 Current Version/draft.tex
reorganize

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={A White Noise Approach to Evolutionary Ecology},
            pdfauthor={Bob Week, Steve Krone, Scott L. Nuismer, Luke J. Harmon},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{A White Noise Approach to Evolutionary Ecology}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Bob Week, Steve Krone, Scott L. Nuismer, Luke J. Harmon}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{3/1/2020}

\usepackage{float}
\usepackage{flafter}

\usepackage{xr}
\usepackage{mathtools}
\usepackage{mathrsfs}
\usepackage{hyperref}
\usepackage[left]{lineno}
\usepackage{amsmath}
\linenumbers
\externaldocument{SM}

>>>>

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={A White Noise Approach to Evolutionary Ecology},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{xr}
\usepackage{mathtools}
\usepackage{mathrsfs}
\usepackage{hyperref}
\usepackage[left]{lineno}
\usepackage{amsmath}
\usepackage[authoryear,sectionbib,sort]{natbib}
\usepackage{fullpage}
\linenumbers
\externaldocument{SM}
\usepackage{float}
\usepackage{flafter}

\title{A White Noise Approach to Evolutionary Ecology}
\author{true \and true \and true \and true}
\date{2020-04-21}
